Streaming CityJSON datasets

### 2024-04-19 paper/main.tex
Try to improve unclear sentences

CityJSON is a JSON-based encoding for storing 3D city models that implements a subset of the latest version, v3.0.0, of the CityGML data model~\citep{OGC-CityGML3}.
As further described in \citet{19_ogdss_cityjson}, its development started in 2017 with the aim of offering an easy-to-use and web-ready alternative to the GML-encoded CityGML files~\citep{OGC-CityGML3-XML}, which in practice can be rather verbose, difficult to parse, and complex to manipulate.
The first official release of CityJSON (version 1.0) was a success: 
\begin{enumerate}
  \item its JSON-based files were on average around 7X more compact than their CityGML-GML equivalents without loss of information. See \citet{19_ogdss_cityjson} and \url{https://www.cityjson.org/filesize/}, and also \citet{Praschl23} for a comparison with standard computer graphics formats; 
  \item it was adopted as an OGC Community Standard~\citep{OGC-CityJSON-v10}; 
  \item it was adopted by the Dutch government as a 3D standard to distribute nation-wide 3D datasets; 
  \item several implementations, \eg\ FME, were developed.
\end{enumerate}

%@HUGO several implementations or plugins?

@@

For GML-based formats (which are feature-centric), modifying a file for the purpose of streaming is usually a simple task that involves sending the features in the dataset one-by-one.
However, notice that this is only true for simple files and GML features stored following the Simple Feature paradigm~\citep{OGC-SF}.
For more complex data models like CityGML where \emph{Implicit Geometries} and \emph{XLinks} are used (for reusing geometries and for storing some topological relationships), the conversion often needs a large amount of pre-processing.
The presence of \emph{XLinks} means that references between elements are possible (\eg\ reusing surfaces), but if in the stream the referenced surface has not appeared, the receiver cannot process the information.
It is not always possible to order CityGML files so that all the references can be resolved without having to store locally information until it appears in the stream.
It is always possible to resolve the \emph{XLinks} before streaming a dataset, but this means that its size is increased, and that converting back to the original file will not be possible.
%@HUGO: I don't understand the last 2 sentenses.

@@

First observe that---contrary to intuition---the size of a dataset, serialised to a JSON file, stored as a CityJSONSeq is around 12\% more compact than as a CityJSON, and in the case of \textbf{3DBAG} it is 36\% compacter.
The main reason for this is that the indices of the vertices are low integers for each feature (because the lowest index in each feature is always ``0''), and no not increase to several millions.
%@Hugo: Sentence above not super clear. 
For instance, the dataset \textbf{Helsinki} contains more than 3 millions vertices, but its largest feature contains only but 2202.
As a result, the CityJSONSeq stored in a file is only but 72\% of the original CityJSON file.
Only one dataset sees its size slightly increased when serialised to a CityJSONSeq file (\textbf{Rotterdam}).
This is caused, we believe, by the fact that the dataset has textures that are shared by most of the buildings, and the \texttt{"appearance"} property must be repeated for each of the feature (in the CityJSON file it is only stored once and referenced by all buildings).

@@

\begin{itemize}
  \item CityJSONSeq is an attractive alternative to CityJSON, in most uses it is better to use since it is compacter, processing it uses less memory and is faster.
  \item We have kept most of the structure of CityJSON so that parsers do not need to be adapted much.
  \item it opens up many possibilities for processing large 3dcm by using Unix pipes. A few of the tools developed for CityJSON are already reading/writing to stdin/stdout, so they can be chained together.
  \item cjseqview, cjseqval are two of them, but developing others to eg calculate volumes or repair geometries would be trivial
  \item the nice part is that they tools do not need to be written in the same language, one pipeline can have tools built in Python and Rust and C++ and it all works nicely.
  \item extensions: these should be placed in the metadata or collection
\end{itemize}

Disadvantages:
\begin{itemize}
  \item     Software implementing CityJSON need to account for both regular CityJSON and CityJSONSeq.
  \item  In case of many, small, interconnecting CityObjects, a global vertex list would provide better compression than the separate vertex lists of the CityJSONFeatures.
  \item  Appearances (esp. Material) are relatively heavy objects with many properties. If an appearance is shared across most CityObjects, duplicating it in each Feature could negate the effect of small vertex indices and actually increase the file size compared to a regular CityJSON file. This is just theorizing, haven't tested.
\end{itemize}

>>>>

CityJSON is a JSON-based encoding for storing 3D city models that implements a subset of the latest version, v3.0.0, of the CityGML data model~\citep{OGC-CityGML3}.
As further described in \citet{19_ogdss_cityjson}, its development started in 2017 with the aim of offering an easy-to-use and web-ready alternative to the GML-encoded CityGML files~\citep{OGC-CityGML3-XML}, which in practice can be rather verbose, difficult to parse, and complex to manipulate.
The first official release of CityJSON (version 1.0) was a success: 
\begin{enumerate}
  \item its JSON-based files were on average around 7X more compact than their CityGML-GML equivalents without loss of information. See \citet{19_ogdss_cityjson} and \url{https://www.cityjson.org/filesize/}, and also \citet{Praschl23} for a comparison with standard computer graphics formats; 
  \item it was adopted as an OGC Community Standard~\citep{OGC-CityJSON-v10}; 
  \item it was adopted by the Dutch government as a 3D standard to distribute nation-wide 3D datasets; 
  \item several implementations and plugins have been developed, most notably FME\@.
\end{enumerate}

@@

For GML-based formats (which are feature-centric), modifying a file for the purpose of streaming is usually a simple task that involves sending the features in the dataset one-by-one.
However, notice that this is only true for simple files and GML features stored following the Simple Feature paradigm~\citep{OGC-SF}.
For more complex data models like CityGML where \emph{Implicit Geometries} and \emph{XLinks} are used, the conversion often needs a large amount of pre-processing.

\emph{XLinks} are links links between elements in a file (similar to \emph{pointers} in a computer program), a concrete example in a CityGML file is a cube that lists the geometries of 5 of its surfaces, but its 6th surface is simply a link to another surface somewhere else in the file (which belongs to another building for instance).
If in the stream the linked surface has not appeared yet, the receiver cannot process the cube and needs to wait for that specific surface to appear.
It is not always possible to (re-)order a CityGML file so that all the references can be resolved without having to store locally information until it appears in the stream.
However, it is always possible to resolve the \emph{XLinks} before streaming a dataset, but this means that its size is increased, and that converting back to the original file will not be possible.

@@

First observe that---contrary to intuition---the size of a dataset, serialised to a JSON file, stored as a CityJSONSeq is around 12\% more compact than as a CityJSON, and in the case of \textbf{3DBAG} it is 36\% compacter.
The main reason for this is that the indices of the vertices are low integers for each feature (because the lowest index in each feature is always ``0''), and no not increase to several millions.
For instance, the dataset \textbf{Helsinki} contains more than 3 millions vertices, but its largest feature contains only but 2202.
The fact that many indices are used for geometries (and for textures) means that if large numbers are used then the size of the file will grow.
As a result, the CityJSONSeq stored in a file is only but 72\% of the original CityJSON file.
Only one dataset sees its size slightly increased when serialised to a CityJSONSeq file (\textbf{Rotterdam}).
This is caused, we believe, by the fact that the dataset has textures that are shared by most of the buildings, and the \texttt{"appearance"} property must be repeated for each of the feature (in the CityJSON file it is only stored once and referenced by all buildings).

@@

\begin{itemize}
  \item CityJSONSeq is an attractive alternative to CityJSON; in most uses it is more compact, its processing requires less memory and it is faster.
  \item We have kept most of the structure of CityJSON so that parsers do not need to be adapted much.
  \item it opens up many possibilities for processing large 3dcm by using Unix pipes. A few of the tools developed for CityJSON are already reading/writing to stdin/stdout, so they can be chained together.
  \item cjseqview, cjseqval are two of them, but developing others to eg calculate volumes or repair geometries would be trivial
  \item the nice part is that they tools do not need to be written in the same language, one pipeline can have tools built in Python and Rust and C++ and it all works nicely.
  \item extensions: these should be placed in the metadata or collection
\end{itemize}

Disadvantages:
\begin{itemize}
  \item  Software implementing CityJSON needs to account for both regular CityJSON and CityJSONSeq.
  \item  In case of many, small, interconnecting CityObjects, a global vertex list would provide better compression than the separate vertex lists of the CityJSONFeatures.
  \item  Appearances (esp. Material) are relatively heavy objects with many properties. If an appearance is shared across most CityObjects, duplicating it in each Feature could negate the effect of small vertex indices and actually increase the file size compared to a regular CityJSON file. This is just theorizing, haven't tested.
\end{itemize}
